---
title: "Files formats & Predicting Future Sales"
output:
  html_document: default
  html_notebook: default
---

## Getting data

Download all datasets from  <https://www.kaggle.com/c/competitive-data-science-predict-future-sales/data> into directory `future_sales_data` and uznip it.

## Access to data from spark


```{r}
library(sparklyr)
library(dplyr)
sc <- spark_connect(master = "local")
```

## Parquet and buckets

```{r}
sales_sdf <- spark_read_csv(sc, "sales", "../future_sales_data/sales_train.csv.gz")

sales_sdf %>%
  mutate(dt=to_timestamp(unix_timestamp(date, 'dd.MM.yyyy'))) %>%
  mutate(year=year(dt), month=month(dt)) %>%
  select(-dt) ->
  sales_sdf
```

```{r}
colname <- 'shop_id'
sales_sdf %>%
  group_by_(colname) %>%
  summarise %>%
  summarise(
    n(),
    max(!!rlang::sym(colname), na.rm = TRUE),
    min(!!rlang::sym(colname), na.rm = TRUE)
  ) 
```

```{r}
sales_sdf %>%
  spark_write_parquet(
    "../future_sales_data/sales_train.parquet",
    partition_by = c("shop_id", "year", "month"))
```


```{r}
sales_sdf <- spark_read_parquet(sc, "sales", "../future_sales_data/sales_train.parquet/shop_id\\=0/year\\=2013/month\\=1/")
```

```{r}
sales_sdf <- spark_read_parquet(sc, "sales", "../future_sales_data/sales_train.parquet/shop_id\\={0, 1}/year\\=2013/month\\=1/")
```

### There is no year

```{r}
sales_sdf %>%
 head
```

```{r}
sales_sdf %>%
  mutate(year=2013) %>%
  head
```

### Function to read from multiple sources

```{r}
library(whisker)

read_sale <- function(shop_id, year, month) {
  
  path_template = "../future_sales_data/sales_train.parquet/shop_id\\={{{shop_id}}}/year\\={{{year}}}/month\\={{{month}}}/"
  data = list(
    shop_id=shop_id,
    month=month,
    year=year)
  path <- whisker.render(path_template, data)
  if (dir.exists(gsub("[\\]", "", path))) {
    spark_read_parquet(sc, "sales", path) %>%
      mutate(
        shop_id = as.integer(shop_id),
        year=as.integer(year),
        month=as.integer(month))
  } else {
    NULL
  }
}
sales_sdf <- read_sale(0, 2014, 1)
sales_sdf
```


```{r}
sales_sdf <- read_sales(0, 2000, 1)
sales_sdf
```

```{r}
read_sales <- function(shop_ids, years, months) {
  sdf <- NULL
  for (shop_id in shop_ids) {
    for (year in years) {
        for (month in months) {
          new_sdf <- read_sale(shop_id, year, month)
          if (!is.null(sdf)) {
            if (!is.null(new_sdf)) {
              sdf <- union_all(sdf, new_sdf)
            }
          } else {
            sdf <- new_sdf
          }
        }
    }
  }
  sdf
}
```

```{r}
sales_sdf <- read_sales(c(0, 1), 2013, 1)
sales_sdf %>%
  group_by(shop_id) %>%
  summarise()
```

```{r}
start_time <- Sys.time()
sales_sdf <- read_sales(0:59, c(2013, 2014, 2015), 1:12)
sales_sdf %>%
  group_by(shop_id) %>%
  summarise %>%
  print
end_time <- Sys.time()
end_time - start_time
```

About 5 mins on mac...

```{r}
sales_sdf %>%
 head
```

### Problem
Could you move this function to a separate file `salesData.R`, import it from there and use it here?


## Project 1: Predict shop's future sales (simplified version)

We want predict next month total sales of a given shop.

Steps to do:

1. Data preparation
2. Train-test splitting
3. Training simple linear model
4. Evaluating: implementing MSE & R2 





